\begin{agradecimentos}
  We create an imaginary show inside our heads, a show that involves more frustrations than the hope of reaching our goal. They are daily struggles, daily failures, so many blows that at first we thought we had lost all this time, but actually, it has been taking you to the highest point, to the point of seeing that small failures were all part of the same gear, the conquest. The energy that always moved me was the disbelief in my work, the lack of support, the disloyalty from part of this department, and the fear of not being able to conclude. But that gear was also \textquotedblleft lubricated\textquotedblright~so many times by faith, family, innumerable friends, places, travels, and so many other events that made things a little less painful. And it is for these people that I toast and dedicate this work today. Thank you to look with me this brand new horizon.    
  % Criamos um espetáculo dentro de nossas cabeças, um espetáculo que envolve mais frustações do que a esperança de atingir nosso objetivo. São lutas diárias, falhas diárias, tantas pancadas que, no ínicio, pensamos ter perdido todo esse tempo, mas que na realidade, lhe levava ao ponto mais alto, ao ponto de enxergar que as pequenas falhas eram todas peças da mesma engrenagem, a da conquista. A energia que sempre movimentou-me foi a descrença em meu trabalho, a falta de suporte, a deslealdade e descaso de parte deste departamento, e o medo de não conseguir. Mas essa engrenagem também foi ``lubrificada'' tantas vezes pela fé, família, pelos inúmeros amigos, lugares, viagens e tantos outros acontecimentos que tornaram as coisas um pouco menos penosas. E são por essas pessoas que brindo e dedico esse trabalho hoje. Obrigado por enxergarem comigo meu mais novo horizonte.
  
  During my PhD, I had the chance to experience the guidance, friendship, support, criticism, and love of incredible people. Everyone who shared this energy played key roles in my evolution and what I am today. Above all, God and Nossa Senhora de Aparecida. Faith and science, the most famous paradox. Lord did not give me the thesis or health, but faith on you will always be the fuel of my determination and hope.
  %Durante o período de meu Doutorado, tive a chance de experienciar a orientação, amizade, suporte, crítica e amor de pessoas incríveis. Todos que compartilharam desta energia, tiveram papéis fundamentais em minha evolução e do que sou hoje. Sobretudo, à Deus e à Nossa Senhora da Aparecida. O famoso paradoxo: fé e ciência. O Senhor não me deu a tese ou saúde, mas a fé em ti será sempre o combustível em minha determinação. 
  
  A special thanks to the great sponsors of this work, my parents, Raul and Zélia, and my sister, Claudia. I can not imagine this work, or anything else, being done without your cooperation or presence. Thanks for everything.
  %Um agradecimento especial aos grandes incentivadores e \textquotedblleft patriocinadores emocional\textquotedblright$~$deste trabalho, meus pais, Raul e Zélia, e minha irmã, Cláudia. Não consigo imaginar este trabalho, ou qualquer outra coisa, sendo realizado sem cooperação ou presença de vocês. Obrigado por tudo. 
  
  To the great INPE, which has hosted me since the master's degree in Applied Computing (2010), to its inspiring infrastructure and environment. To the eternal friends of the postgraduate class in Remote Sensing of 2014, with affection to my friends Aline Jacon, Ana Pessoa, Bruna Pechini, Cesare Neto, David França, Denis Mariano, Henrique Cassol, Jaidson Becker, João Felipe , João Bosco, Ranieli dos Anjos, Sacha Siani and all others who were with me. My friends from Hannover, in special Alexander Schunert, Lukas Schack and Joachim Niemeyer. To my dear friends and advisors Dr. Luiz Aragão and Dr. Yosio Shimabukuro, for having taken this work already underway, for believing, for cooperation and friendship until the end. Dr. Elisabete Moraes, my academic mother and great friend. Thank you for embracing the cause so many times, for understanding and for struggling with me. This work is a special dedication to you.
  %À grande instituição INPE, que abrigou-me desde o mestrado em CAP (2010), à sua infraestrutura e ambiente inspiradores. Aos eternos amigos da turma de pós-graduação em Sensoriamento Remoto de 2014, com carinho, aos singulares Sacha Siani, David França, Aline Jacon, Ana Pessoa, Ranieli dos Anjos, Bruna Pechini, Cesare Neto, Henrique Cassol, Jaidson Becker, João Felipe, João Bosco, Dênis Mariano e todos outros que estiveram comigo. Aos caros amigos e orientadores Dr. Luiz Aragão e Dr. Yosio Shimabukuro, por terem assumido este trabalho já em curso, por acreditarem, pela cooperação e pela amizade até o término. Dra. Elisabete Moraes, minha mãe acadêmica e grande amiga. Obrigado por abraçar a causa tantas vezes, pela compreensão e por ter lutado comigo. Este trabalho é uma dedicatória especial à você. 
  
  To the dear friends of the Institute of Photogrammetry (IfP), University of Stuttgart, Germany: Gottfried Mandlburger, Luka Jurjevic, Mateusz Karpina, Patrick Tutzauer, Stefan Schmohl, Dominik Laupheimer, Martina Kroma, Lavinia Runceanu, Chia-Hsiang, Ke Gong, Prof. Dr. Michael Cramer and, in particular, Prof. Dr. Uwe Sörgel and Prof. Dr. Norbert Haala, for believing in my work, for the support, reception and follow-up of all my work during the time I was in Stuttgart.
  %Aos queridos amigos do Departamento de Fotogrametria da Universidade de Stuttgart (IfP), Alemanha, Gottfried Mandlburger, Luka Jurjevic, Mateusz Karpina, Patrick Tutzauer, Stefan Schmohl, Dominik Laupheimer, Martina Kroma, Lavinia Runceanu, Chia-Hsiang, Ke Gong, Prof. Dr. Michael Cramer e, em especial, Prof. Dr. Uwe Sörgel e Dr. Prof. Norbert Haala, por acreditarem em meu trabalho, pelo suporte, recepção e acompanhamento de todo meu trabalho durante o período em que estive em Stuttgart. 
  
  My co-work friends, at Brazilian Space Weather Monitoring Program (EMBRACE/INPE): Dr. Cristiano Wrasse, Dr. Clezio De Nardin, Fauéz and Débora, for the enormous collaboration, infrastructure and support in this journey. Undoubtedly, the valuable lessons and experience as an Analyst and Developer in this department have brought great professional projections and a new insight into practical aspects in problem solving.
  %Aos amigos de trabalho, no Programa de Monitoramento Brasileiro de Clima Espacial (EMBRACE/INPE), Dr. Cristiano Wrasse, Dr. Clezio De Nardin, Fauéz e Débora, pela enorme colaboração, infraestrutura e apoio nesta jornada. Sem dúvida, os valiosos ensinamentos e experiência como Analista e Desenvolvedor neste departamento trouxeram ótimas projeções profissionais e uma nova percepção sobre aspectos práticos na resolução de problemas.
  
  Finally, to the Coordination for the Improvement of Higher Education Personnel (CAPES) and the National Council for Scientific and Technological Development (CNPq), thank you for the financial support in my first year of doctorate, and for the great opportunity and teachings I had at the University of Stuttgart (IfP-Stuttgart) - grant PDSE, Process No. 88881.132115/2016-01. This great opportunity was undoubtedly a very important piece in the conclusion of this work. The Army Geographic Service Directorate (DSG), for their support for cartographic specifications regarding 3D mapping in Brazil. To the friend George Longhitano, G-Drones, for the cooperation in acquiring and providing the data for our experiments.
  %Por fim, à Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) e ao Conselho Nacional de Desenvolvimento Científico e Tecnológico (CNPq), obrigado pelo suporte financeiro em meu primeiro ano de doutorado, e pela grande oportunidade e ensinamentos que tive na Universidade de Stuttgart (IfP-Stuttgart). Essa grande oportunidade foi, sem dúvida, peça importantíssima na conclusão deste trabalho.
  
  \textit{Nobody inside this institute was a motivation for me more than myself. No one will motivate you more than your own determination and courage. This was the biggest win of my life, which will be small compared to my brand new challenges.}
  
\end{agradecimentos}


